---
title: "17_explore_results_cell_permutations"
author: "Rogério Ribeiro"
date: "`r Sys.Date()`"
output: html_document
---

# Explore Results from the scRNA differential variability analysis

# Load libraries
```{r}

library(tidyverse)
library(ggrastr)
library(reshape2)
library(ComplexHeatmap)
library(RColorBrewer)
library(gridExtra)
library(ggrepel)
library(ggpubr)
```


# Set image dir
```{r}
img.dir <- "../../results/azimuth_annotation_results/permutation_results_exploration/"
if (!dir.exists(img.dir)) {dir.create(img.dir)}
```



# Load data
```{r}
cts <- c(
  "Bintermediate", "Bmemory", "Bnaive", 
  "CD14Mono", "CD16Mono", "CD4CTL", "CD4Naive", 
  "CD4Proliferating", "CD4TCM", "CD4TEM", 
  "CD8Naive", "CD8Proliferating", "CD8TCM", "CD8TEM", 
  "Eryth", "HSPC", "MAIT", 
  "NKProliferating", "NK_CD56bright", "NK", 
  "Plasmablast", "Platelet", "Treg", 
  "cDC2", "dnT", "gdT", "pDC"
)

cts_to_study <- c("Bintermediate", "Bmemory", "Bnaive", "CD4CTL", "CD4Naive", "CD4TCM", "CD4TEM", "CD8Naive", 
                "CD8TCM", "CD8TEM", "CD14Mono", "CD16Mono","MAIT", "NK_CD56bright", "NK", "Treg")

```


## Load mean expression


```{r}
mean_exps <- list()
for (ct in cts){
  mean_exps[[ct]] <- read.table(paste0("../../results/azimuth_annotation_results/03_mean_variance_estimates/", ct, "_cells_mean_mx.txt"))
}

```

## Load dispersion estimates

```{r}
dispersions <- list()
for (ct in cts){
  dispersions[[ct]] <- read.table(paste0("../../results/azimuth_annotation_results/04_individual_dispersions/", ct, "_within_ind_dispersion.txt"), header = TRUE)
  row.names(dispersions[[ct]]) <- dispersions[[ct]]$gene
  dispersions[[ct]]$gene <- NULL
}
```

## Load DVG list

```{r}
list_of_permutations <- list()
for (i in 1:10){
  DVGs <- readRDS(paste0("../../results/azimuth_annotation_results/17_model_fit_combined/", i, ".DVG_results.rds")) # Open file
  DVGs <- lapply(DVGs, function(x) lapply(x, function(y) y %>% mutate(significance=ifelse(p_value_adj < 0.05, "sig", "not_sig"))))
  
  # Remove platelets and Erythrocytes
  DVGs$age$Platelets <- NULL
  DVGs$sex$Platelets <- NULL
  
  DVGs$age$Erythrocytes <- NULL
  DVGs$sex$Erythrocytes <- NULL  
  list_of_permutations[[i]] <- DVGs
}
```

# Load Analysis results
```{r}
DVGs <- readRDS("../../results/azimuth_annotation_results/06_model_fit/1.DVG_results.rds") # 
DVGs <- lapply(DVGs, function(x) lapply(x, function(y) y %>% mutate(significance=ifelse(p_value_adj < 0.05, "sig", "not_sig"))))
  
# Remove platelets and Erythrocytes
DVGs$age$Platelets <- NULL
DVGs$sex$Platelets <- NULL

DVGs$age$Erythrocytes <- NULL
DVGs$sex$Erythrocytes <- NULL  
```



# Before filtering
## Count the number of differential expressed genes in each iteration for each celltype
```{r}
# age
number_genes_age_across_permutations <- list()
for (i in 1:10){
  results <- list_of_permutations[[i]]
  number_of_genes_age <- lapply(names(results$age), function(x){
      temp <- results$age[[x]] %>% filter(significance == "sig")
      data.frame("celltype" = x, "genes" = nrow(temp), "iter" = i)
    }
  )

  number_genes_age_across_permutations[[i]] <- do.call("rbind", number_of_genes_age)
}

number_genes_age_across_permutations <- do.call("rbind",number_genes_age_across_permutations)


number_of_genes_age_analysis <- lapply(names(DVGs$age), function(x){
      temp <- DVGs$age[[x]] %>% filter(significance == "sig")
      data.frame("celltype" = x, "genes" = nrow(temp), "iter" = "Analysis")
  }
)

number_of_genes_age_analysis <- do.call("rbind", number_of_genes_age_analysis)

number_genes_age_across_permutations <- rbind(number_genes_age_across_permutations, number_of_genes_age_analysis)

number_genes_age_across_permutations$type <- ifelse(number_genes_age_across_permutations$iter ==  "Analysis", "real", "simulation")


# Sex
number_genes_sex_across_permutations <- list()
for (i in 1:10){
  results <- list_of_permutations[[i]]
  number_of_genes_sex <- lapply(names(results$sex), function(x){
      temp <- results$sex[[x]] %>% filter(significance == "sig")
      data.frame("celltype" = x, "genes" = nrow(temp), "iter" = i)
    }
  )

  number_genes_sex_across_permutations[[i]] <- do.call("rbind", number_of_genes_sex)
}

number_genes_sex_across_permutations <- do.call("rbind",number_genes_sex_across_permutations)


number_of_genes_sex_analysis <- lapply(names(DVGs$sex), function(x){
      temp <- DVGs$sex[[x]] %>% filter(significance == "sig")
      data.frame("celltype" = x, "genes" = nrow(temp), "iter" = "Analysis")
  }
)

number_of_genes_sex_analysis <- do.call("rbind", number_of_genes_sex_analysis)

number_genes_sex_across_permutations <- rbind(number_genes_sex_across_permutations, number_of_genes_sex_analysis)

number_genes_sex_across_permutations$type <- ifelse(number_genes_sex_across_permutations$iter == "Analysis", "real", "simulation")

```

## Plot permutation vs ground truth

```{r fig.height=6, fig.width=10}
number_genes_age_across_permutations.h <- number_genes_age_across_permutations %>% filter(celltype %in% c("CD4_NC", "NK", "CD8_ET", "CD8_NC", "B_IN", "B_MEM", "CD4_ET", "CD8_S100B"))
number_genes_age_across_permutations.l <- number_genes_age_across_permutations %>% filter(!celltype %in% c("CD4_NC", "NK", "CD8_ET", "CD8_NC", "B_IN", "B_MEM", "CD4_ET", "CD8_S100B"))

a <- ggplot(data = number_genes_age_across_permutations.h, aes(x = celltype, y = genes, colour = type)) +
  geom_jitter(width = 0.2, size = 3, alpha = 0.7) +
  labs(
    title = "Permutation vs Real before filtering (age)",
    color = " "
  ) +
  scale_colour_manual(values = c("simulation" = "blue", "real" = "darkred")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "right",
    axis.text = element_text(angle = 90, vjust = 1)
  ) +
  xlab("") + ylab("")

b <- ggplot(data = number_genes_age_across_permutations.l, aes(x = celltype, y = genes, colour = type)) +
  geom_jitter(width = 0.2, size = 3, alpha = 0.7) +
  labs(
    title = "Permutation vs Real before filtering (age)",
    color = " "
  ) +
  scale_colour_manual(values = c("simulation" = "blue", "real" = "darkred")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "right",
    axis.text = element_text(angle = 90, vjust = 1)
  ) +
  xlab("") + ylab("")


number_genes_sex_across_permutations.h <- number_genes_sex_across_permutations %>% filter(celltype %in% c("CD4_NC","CD8_NC"))
number_genes_sex_across_permutations.l <- number_genes_sex_across_permutations %>% filter(!celltype %in% c("CD4_NC","CD8_NC"))

c <- ggplot(data = number_genes_sex_across_permutations.h, aes(x = celltype, y = genes, colour = type)) +
  geom_jitter(width = 0.2, size = 3, alpha = 0.7) +
  labs(
    title = "Permutation vs Real before filtering (sex)",
    color = " "
  ) +
  scale_colour_manual(values = c("simulation" = "blue", "real" = "darkred")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "right",
    axis.text = element_text(angle = 90, vjust = 1)
  ) +
  xlab("") + ylab("")

d <- ggplot(data = number_genes_sex_across_permutations.l, aes(x = celltype, y = genes, colour = type)) +
  geom_jitter(width = 0.2, size = 3, alpha = 0.7) +
  labs(
    title = "Permutation vs Real before filtering (sex)",
    color = " "
  ) +
  scale_colour_manual(values = c("simulation" = "blue", "real" = "darkred")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "right",
    axis.text = element_text(angle = 90, vjust = 1)
  ) +
  xlab("") + ylab("")

bef_filter <- ggarrange(a,b,c,d, ncol = 2, nrow = 2, common.legend = TRUE)


pdf(paste0(img.dir, "permutation_analysis_before_filtering.pdf"), w  = 8, h = 6)
bef_filter
dev.off()

bef_filter
```

# Apply the expression filter to the data---- 

```{r}
# This is specific for the celltype. In the paper they used 0.3 for CD4nc. Here I will use 0.3 for CD4 and other celltypes

filter_mean_ct <- c(10, 10, 5,30,1.5,2,10,5,30,2,10,30,30,30,2,10) #Manually selected based on the simulation
names(filter_mean_ct) <- cts_to_study
```


## Filter using celltype specific filters
```{r}
#adjusted.pvalues are recomputed

number_genes_age_across_permutations.filtered <- lapply(names(filter_mean_ct), function(x){
  mean_per_gene <- rowMeans(mean_exps[[x]], na.rm = TRUE)
  th <- filter_mean_ct[x]
  genes_to_keep <- names(mean_per_gene[mean_per_gene >= th])
  genes_in_permutation_after_filtering <- list()
  for (i in 1:10){
     DVG_cell <- list_of_permutations[[i]]$age[[x]]
     DVG_cell <- DVG_cell %>% 
        filter(gene %in% genes_to_keep) %>% 
        mutate(p_value_adj_filter = p.adjust(p_value, method = "BH"),
               significance_after_filter = ifelse(p_value_adj_filter < 0.05, "sig", "non_sig"))
     genes_in_permutation_after_filtering[[i]] <- data.frame(ct = x, nGene = sum(DVG_cell$p_value_adj_filter < 0.05), iter = i)
  }
  do.call("rbind", genes_in_permutation_after_filtering)
})

number_genes_sex_across_permutations.filtered <- lapply(names(filter_mean_ct), function(x){
  mean_per_gene <- rowMeans(mean_exps[[x]], na.rm = TRUE)
  th <- filter_mean_ct[x]
  genes_to_keep <- names(mean_per_gene[mean_per_gene >= th])
  genes_in_permutation_after_filtering <- list()
  for (i in 1:10){
     DVG_cell <- list_of_permutations[[i]]$sex[[x]]
     DVG_cell <- DVG_cell %>% 
        filter(gene %in% genes_to_keep) %>% 
        mutate(p_value_adj_filter = p.adjust(p_value, method = "BH"),
               significance_after_filter = ifelse(p_value_adj_filter < 0.05, "sig", "non_sig"))
     genes_in_permutation_after_filtering[[i]] <- data.frame(ct = x, nGene = sum(DVG_cell$p_value_adj_filter < 0.05), "iter" = i)
  }
  do.call("rbind", genes_in_permutation_after_filtering)
})


number_genes_age_across_permutations.filtered <- do.call("rbind", number_genes_age_across_permutations.filtered)
number_genes_sex_across_permutations.filtered <- do.call("rbind", number_genes_sex_across_permutations.filtered)


# Load real data and add it to the table
DVG.filter <- readRDS("../../results/azimuth_annotation_results/09_explore_DVA_results/1.DVG.group.filter.rds")

number_of_genes_age.filter <- lapply(names(DVG.filter$age), function(x){
    temp <- DVG.filter$age[[x]] %>% filter(significance_after_filter == "sig")
    data.frame("ct" = x, "nGene" = nrow(temp), iter = "Real")
  }
)

number_of_genes_age.filter <- do.call("rbind", number_of_genes_age.filter)
number_genes_age_across_permutations.filtered <- rbind(number_genes_age_across_permutations.filtered, number_of_genes_age.filter)

# sex 
number_of_genes_sex.filter <- lapply(names(DVG.filter$sex), function(x){
    temp <- DVG.filter$sex[[x]] %>% filter(significance_after_filter == "sig")
    data.frame("ct" = x, "nGene" = nrow(temp), iter = "Real")
  }
)

number_of_genes_sex.filter <- do.call("rbind", number_of_genes_sex.filter)
number_genes_sex_across_permutations.filtered <- rbind(number_genes_sex_across_permutations.filtered, number_of_genes_sex.filter)

```



## Plot permutation vs ground truth

```{r fig.height=6, fig.width=10}

number_genes_age_across_permutations.filtered$type <- ifelse(number_genes_age_across_permutations.filtered$iter == "Real", "real", "simulation") 

number_genes_age_across_permutations.filtered.h <- number_genes_age_across_permutations.filtered %>% filter(ct %in% c("CD4_NC", "NK", "CD8_ET", "CD8_NC"))
number_genes_age_across_permutations.filtered.l <- number_genes_age_across_permutations.filtered %>% filter(!ct %in% c("CD4_NC", "NK", "CD8_ET", "CD8_NC"))

a <- ggplot(data = number_genes_age_across_permutations.filtered.h, aes(x = ct, y = nGene, colour = type)) +
  geom_jitter(width = 0.2, size = 3, alpha = 0.7) +
  labs(
    title = "Permutation vs Real before filtering (Age)",
    color = " "
  ) +
  scale_colour_manual(values = c("simulation" = "blue", "real" = "darkred")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "right",
    axis.text = element_text(angle = 90, vjust = 1)
  ) +
  xlab("") + ylab("")

b <- ggplot(data = number_genes_age_across_permutations.filtered.l, aes(x = ct, y = nGene, colour = type)) +
  geom_jitter(width = 0.2, size = 3, alpha = 0.7) +
  labs(
    title = "Permutation vs Real before filtering (age)",
    color = " "
  ) +
  scale_colour_manual(values = c("simulation" = "blue", "real" = "darkred")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "right",
    axis.text = element_text(angle = 90, vjust = 1)
  ) +
  xlab("") + ylab("")


number_genes_sex_across_permutations.filtered$type <- ifelse(number_genes_sex_across_permutations.filtered$iter == "Real", "real", "simulation") 

number_genes_sex_across_permutations.filtered.h <- number_genes_sex_across_permutations.filtered %>% filter(ct %in% c("B_IN", "CD4_NC", "NK", "CD8_ET"))
number_genes_sex_across_permutations.filtered.l <- number_genes_sex_across_permutations.filtered %>% filter(!ct %in% c("B_IN","CD4_NC", "NK", "CD8_ET"))


c <- ggplot(data = number_genes_sex_across_permutations.filtered.h, aes(x = ct, y = nGene, colour = type)) +
  geom_jitter(width = 0.2, size = 3, alpha = 0.7) +
  labs(
    title = "Permutation vs Real before filtering (sex)",
    color = " "
  ) +
  scale_colour_manual(values = c("simulation" = "blue", "real" = "darkred")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "right",
    axis.text = element_text(angle = 90, vjust = 1)
  ) +
  xlab("") + ylab("")

d <- ggplot(data = number_genes_sex_across_permutations.filtered.l, aes(x = ct, y = nGene, colour = type)) +
  geom_jitter(width = 0.2, size = 3, alpha = 0.7) +
  labs(
    title = "Permutation vs Real before filtering (sex)",
    color = " "
  ) +
  scale_colour_manual(values = c("simulation" = "blue", "real" = "darkred")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "right",
    axis.text = element_text(angle = 90, vjust = 1)
  ) +
  xlab("") + ylab("")




after <- ggarrange(a,b,c,d, ncol = 2, nrow = 2, common.legend = TRUE)


pdf(paste0(img.dir, "permutation_analysis_after_filtering.pdf"), w  = 8, h = 6)
after
dev.off()

after
```
