---
title: "09_enrichement_results"
author: "Rogério Ribeiro"
date: "`r Sys.Date()`"
output: html_document
---

# Load libraries

```{r}
suppressPackageStartupMessages(library(tidyverse))
library(ggpubr)
library(DOSE)
```

# Make out dir

```{r}
#out.dir <- "../../results/azimuth_annotation_results/_enrichement/"
```

# Load data

```{r}
enrichement_age <- readRDS("../../results/azimuth_annotation_results/_enrichement/enrichements_age.rds")
enrichement_sex <- readRDS("../../results/azimuth_annotation_results/_enrichement/enrichements_sex.rds")

enrichement_M <- readRDS("../../results/azimuth_annotation_results/_enrichement/enrichements_age_stratified_M.rds")
enrichement_Fe <- readRDS("../../results/azimuth_annotation_results/_enrichement/enrichements_age_stratified_Fe.rds")


```

# Set pallete

```{r}
safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")
```

# Summarise the results across celltypes

Set function

```{r}
summarise_enrichement <- function(object, ont, FDR){
  if (!ont %in% names(object)){
    return (NULL)
  }
  data <- object[[ont]]@result
  data$ontology <- ont
  data %>% filter(p.adjust < FDR)
}
```

## Age

```{r}
# Summarise BP (Biological Process) enrichment
age.bp.up <- do.call("rbind", lapply(enrichement_age, function(x) summarise_enrichement(x, "BP_up", FDR = 0.05)))
age.bp.down <- do.call("rbind", lapply(enrichement_age, function(x) summarise_enrichement(x, "BP_down", FDR = 0.05)))
age.bp <- rbind(age.bp.up, age.bp.down)

# Summarise MF (Molecular Function) enrichment
age.mf.up <- do.call("rbind", lapply(enrichement_age, function(x) summarise_enrichement(x, "MF_up", FDR = 0.05)))
age.mf.down <- do.call("rbind", lapply(enrichement_age, function(x) summarise_enrichement(x, "MF_down", FDR = 0.05)))
age.mf <- rbind(age.mf.up, age.mf.down)

# Summarise CC (Cellular Component) enrichment
age.cc.up <- do.call("rbind", lapply(enrichement_age, function(x) summarise_enrichement(x, "CC_up", FDR = 0.05)))
age.cc.down <- do.call("rbind", lapply(enrichement_age, function(x) summarise_enrichement(x, "CC_down", FDR = 0.05)))
age.cc <- rbind(age.cc.up, age.cc.down)

# Summarise KEGG pathway enrichment
age.kegg.up <- do.call("rbind", lapply(enrichement_age, function(x) summarise_enrichement(x, "kegg_up", FDR = 0.05)))
age.kegg.down <- do.call("rbind", lapply(enrichement_age, function(x) summarise_enrichement(x, "kegg_down", FDR = 0.05)))
age.kegg <- rbind(age.kegg.up, age.kegg.down)

# Summarise DO (Disease Ontology) enrichment
age.do.up <- do.call("rbind", lapply(enrichement_age, function(x) summarise_enrichement(x, "do_up", FDR = 0.05)))
age.do.down <- do.call("rbind", lapply(enrichement_age, function(x) summarise_enrichement(x, "do_down", FDR = 0.05)))
age.do <- rbind(age.do.up, age.do.down)

# Summarise DNG (Drug Ontology) enrichment
age.dng.up <- do.call("rbind", lapply(enrichement_age, function(x) summarise_enrichement(x, "dng_up", FDR = 0.05)))
age.dng.down <- do.call("rbind", lapply(enrichement_age, function(x) summarise_enrichement(x, "dng_down", FDR = 0.05)))
age.dng <- rbind(age.dng.up, age.dng.down)
```

## Sex

```{r}
# Summarise BP (Biological Process) enrichment
sex.bp.up <- do.call("rbind", lapply(enrichement_sex, function(x) summarise_enrichement(x, "BP_up", FDR = 0.05)))
sex.bp.down <- do.call("rbind", lapply(enrichement_sex, function(x) summarise_enrichement(x, "BP_down", FDR = 0.05)))
sex.bp <- rbind(sex.bp.up, sex.bp.down)

# Summarise MF (Molecular Function) enrichment
sex.mf.up <- do.call("rbind", lapply(enrichement_sex, function(x) summarise_enrichement(x, "MF_up", FDR = 0.05)))
sex.mf.down <- do.call("rbind", lapply(enrichement_sex, function(x) summarise_enrichement(x, "MF_down", FDR = 0.05)))
sex.mf <- rbind(sex.mf.up, sex.mf.down)

# Summarise CC (Cellular Component) enrichment
sex.cc.up <- do.call("rbind", lapply(enrichement_sex, function(x) summarise_enrichement(x, "CC_up", FDR = 0.05)))
sex.cc.down <- do.call("rbind", lapply(enrichement_sex, function(x) summarise_enrichement(x, "CC_down", FDR = 0.05)))
sex.cc <- rbind(sex.cc.up, sex.cc.down)

# Summarise KEGG pathway enrichment
sex.kegg.up <- do.call("rbind", lapply(enrichement_sex, function(x) summarise_enrichement(x, "kegg_up", FDR = 0.05)))
sex.kegg.down <- do.call("rbind", lapply(enrichement_sex, function(x) summarise_enrichement(x, "kegg_down", FDR = 0.05)))
sex.kegg <- rbind(sex.kegg.up, sex.kegg.down)

# Summarise DO (Disease Ontology) enrichment
sex.do.up <- do.call("rbind", lapply(enrichement_sex, function(x) summarise_enrichement(x, "do_up", FDR = 0.05)))
sex.do.down <- do.call("rbind", lapply(enrichement_sex, function(x) summarise_enrichement(x, "do_down", FDR = 0.05)))
sex.do <- rbind(sex.do.up, sex.do.down)

# Summarise DNG (Drug Ontology) enrichment
sex.dng.up <- do.call("rbind", lapply(enrichement_sex, function(x) summarise_enrichement(x, "dng_up", FDR = 0.05)))
sex.dng.down <- do.call("rbind", lapply(enrichement_sex, function(x) summarise_enrichement(x, "dng_down", FDR = 0.05)))
sex.dng <- rbind(sex.dng.up, sex.dng.down)
```

# Summarise the results with a table

## Age

```{r fig.height=10, fig.width=15}
# Biological process 
celltypes <- do.call("c", lapply(row.names(age.bp), function(x) str_split(x, "\\.")[[1]][1]))
age.bp$ct <- celltypes

p1 <- ggplot(age.bp, aes(y = ct, x = Description, color = ontology)) +
  geom_point(size = 4) +
  scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
  theme_dose(10) +
  xlab("") + ylab("") + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
        axis.text.y = element_text(size = 16), 
        legend.title = element_blank(), 
        legend.text = element_text(size = 22), 
        legend.position = "top", 
        legend.direction="horizontal")

# MOlecular function
celltypes <- do.call("c", lapply(row.names(age.mf), function(x) str_split(x, "\\.")[[1]][1]))
age.mf$ct <- celltypes

p2 <- ggplot(age.mf, aes(y = ct, x = Description, color = ontology)) +
  geom_point(size = 4) +
  scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
  theme_dose(10) +
  xlab("") + ylab("") + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
        axis.text.y = element_text(size = 16), 
        legend.title = element_blank(), 
        legend.text = element_text(size = 22), 
        legend.position = "top", 
        legend.direction="horizontal")


# Cellular component 
celltypes <- do.call("c", lapply(row.names(age.cc), function(x) str_split(x, "\\.")[[1]][1]))
age.cc$ct <- celltypes

p3 <- ggplot(age.cc, aes(y = ct, x = Description, color = ontology)) +
  geom_point(size = 4) +
  scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
  theme_dose(10) +
  xlab("") + ylab("") + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
        axis.text.y = element_text(size = 16), 
        legend.title = element_blank(), 
        legend.text = element_text(size = 22), 
        legend.position = "top", 
        legend.direction="horizontal")

# KEGG
celltypes <- do.call("c", lapply(row.names(age.kegg), function(x) str_split(x, "\\.")[[1]][1]))
age.kegg$ct <- celltypes

p4 <- ggplot(age.kegg, aes(y = ct, x = Description, color = ontology)) +
  geom_point(size = 4) +
  scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
  theme_dose(10) +
  xlab("") + ylab("") + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
        axis.text.y = element_text(size = 16), 
        legend.title = element_blank(), 
        legend.text = element_text(size = 22), 
        legend.position = "top", 
        legend.direction="horizontal")
# 
# # Disease Ontology
# celltypes <- do.call("c", lapply(row.names(age.do), function(x) str_split(x, "\\.")[[1]][1]))
# age.do$ct <- celltypes
# 
# p5 <- ggplot(age.do, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") + 
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
#         axis.text.y = element_text(size = 16), 
#         legend.title = element_blank(), 
#         legend.text = element_text(size = 22), 
#         legend.position = "top", 
#         legend.direction="horizontal")

# Disease Network
celltypes <- do.call("c", lapply(row.names(age.dng), function(x) str_split(x, "\\.")[[1]][1]))
age.dng$ct <- celltypes

p6 <- ggplot(age.dng, aes(y = ct, x = Description, color = ontology)) +
  geom_point(size = 4) +
  scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
  theme_dose(10) +
  xlab("") + ylab("") + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
        axis.text.y = element_text(size = 16), 
        legend.title = element_blank(), 
        legend.text = element_text(size = 22), 
        legend.position = "top", 
        legend.direction="horizontal")




print(p1)
print(p2)
print(p3)
print(p4)
#print(p5)
print(p6)
```

```{r}

pdf("../../results/azimuth_annotation_results/_enrichement/images/bp.enrich.age.pdf")
p1
dev.off()


pdf("../../results//azimuth_annotation_results/_enrichement/images/mf.enrich.age.pdf", h = 10)
p2
dev.off()

pdf("../../results/azimuth_annotation_results/_enrichement//images/cc.enrich.age.pdf", w = 8)
p3
dev.off()

pdf("../../results/azimuth_annotation_results/_enrichement//images/KEGG.enrich.age.pdf", w = 8)
p4
dev.off()

pdf("../../results/azimuth_annotation_results/_enrichement//images/DNG.enrich.age.pdf", w = 16, h = 12)
p6
dev.off()

```

## Sex

```{r fig.height=10, fig.width=15}
# Biological process 
celltypes <- do.call("c", lapply(row.names(sex.bp), function(x) str_split(x, "\\.")[[1]][1]))
sex.bp$ct <- celltypes
# 
# p1 <- ggplot(sex.bp, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") + 
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
#         axis.text.y = element_text(size = 16), 
#         legend.title = element_blank(), 
#         legend.text = element_text(size = 22), 
#         legend.position = "top", 
#         legend.direction="horizontal")

# MOlecular function
celltypes <- do.call("c", lapply(row.names(sex.mf), function(x) str_split(x, "\\.")[[1]][1]))
sex.mf$ct <- celltypes

p2 <- ggplot(sex.mf, aes(y = ct, x = Description, color = ontology)) +
  geom_point(size = 4) +
  scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
  theme_dose(10) +
  xlab("") + ylab("") + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
        axis.text.y = element_text(size = 16), 
        legend.title = element_blank(), 
        legend.text = element_text(size = 22), 
        legend.position = "top", 
        legend.direction="horizontal")


# Cellular component 
celltypes <- do.call("c", lapply(row.names(sex.cc), function(x) str_split(x, "\\.")[[1]][1]))
sex.cc$ct <- celltypes

p3 <- ggplot(sex.cc, aes(y = ct, x = Description, color = ontology)) +
  geom_point(size = 4) +
  scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
  theme_dose(10) +
  xlab("") + ylab("") + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
        axis.text.y = element_text(size = 16), 
        legend.title = element_blank(), 
        legend.text = element_text(size = 22), 
        legend.position = "top", 
        legend.direction="horizontal")

# KEGG
celltypes <- do.call("c", lapply(row.names(sex.kegg), function(x) str_split(x, "\\.")[[1]][1]))
sex.kegg$ct <- celltypes

p4 <- ggplot(sex.kegg, aes(y = ct, x = Description, color = ontology)) +
  geom_point(size = 4) +
  scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
  theme_dose(10) +
  xlab("") + ylab("") + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
        axis.text.y = element_text(size = 16), 
        legend.title = element_blank(), 
        legend.text = element_text(size = 22), 
        legend.position = "top", 
        legend.direction="horizontal")

# Disease Ontology
celltypes <- do.call("c", lapply(row.names(sex.do), function(x) str_split(x, "\\.")[[1]][1]))
sex.do$ct <- celltypes
# 
# p5 <- ggplot(sex.do, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") + 
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
#         axis.text.y = element_text(size = 16), 
#         legend.title = element_blank(), 
#         legend.text = element_text(size = 22), 
#         legend.position = "top", 
#         legend.direction="horizontal")

# Disease Network
celltypes <- do.call("c", lapply(row.names(sex.dng), function(x) str_split(x, "\\.")[[1]][1]))
sex.dng$ct <- celltypes

# p6 <- ggplot(sex.dng, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") + 
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
#         axis.text.y = element_text(size = 16), 
#         legend.title = element_blank(), 
#         legend.text = element_text(size = 22), 
#         legend.position = "top", 
#         legend.direction="horizontal")




#print(p1)
print(p2)
print(p3)
print(p4)
#print(p5)
#print(p6)
```

```{r}

pdf("../../results/azimuth_annotation_results/_enrichement//images/bp.enrich.sex.pdf")
p1
dev.off()


pdf("../../results/azimuth_annotation_results/_enrichement//images/mf.enrich.sex.pdf", h = 10)
p2
dev.off()

pdf("../../results/azimuth_annotation_results/_enrichement//images/cc.enrich.sex.pdf", w = 8)
p3
dev.off()

pdf("../../results/azimuth_annotation_results/_enrichement//images/KEGG.enrich.sex.pdf", w = 8)
p4
dev.off()

pdf("../../results/azimuth_annotation_results/_enrichement//images/DO.enrich.sex.pdf", w = 8)
#p5
dev.off()

pdf("../../results/azimuth_annotation_results/_enrichement//images/DNG.enrich.sex.pdf", w = 16, h = 12)
p6
dev.off()

```



# Repeat for stratified results in age (by sex)
## Summarise the results across celltypes

## Male

```{r}
# Summarise BP (Biological Process) enrichment
M.bp.up <- do.call("rbind", lapply(enrichement_M, function(x) summarise_enrichement(x, "BP_up", FDR = 0.05)))
M.bp.down <- do.call("rbind", lapply(enrichement_M, function(x) summarise_enrichement(x, "BP_down", FDR = 0.05)))
M.bp <- rbind(M.bp.up, M.bp.down)

# Summarise MF (Molecular Function) enrichment
M.mf.up <- do.call("rbind", lapply(enrichement_M, function(x) summarise_enrichement(x, "MF_up", FDR = 0.05)))
M.mf.down <- do.call("rbind", lapply(enrichement_M, function(x) summarise_enrichement(x, "MF_down", FDR = 0.05)))
M.mf <- rbind(M.mf.up, M.mf.down)

# Summarise CC (Cellular Component) enrichment
M.cc.up <- do.call("rbind", lapply(enrichement_M, function(x) summarise_enrichement(x, "CC_up", FDR = 0.05)))
M.cc.down <- do.call("rbind", lapply(enrichement_M, function(x) summarise_enrichement(x, "CC_down", FDR = 0.05)))
M.cc <- rbind(M.cc.up, M.cc.down)

# Summarise KEGG pathway enrichment
M.kegg.up <- do.call("rbind", lapply(enrichement_M, function(x) summarise_enrichement(x, "kegg_up", FDR = 0.05)))
M.kegg.down <- do.call("rbind", lapply(enrichement_M, function(x) summarise_enrichement(x, "kegg_down", FDR = 0.05)))
M.kegg <- rbind(M.kegg.up, M.kegg.down)

# Summarise DO (Disease Ontology) enrichment
M.do.up <- do.call("rbind", lapply(enrichement_M, function(x) summarise_enrichement(x, "do_up", FDR = 0.05)))
M.do.down <- do.call("rbind", lapply(enrichement_M, function(x) summarise_enrichement(x, "do_down", FDR = 0.05)))
M.do <- rbind(M.do.up, M.do.down)

# Summarise DNG (Drug Ontology) enrichment
M.dng.up <- do.call("rbind", lapply(enrichement_M, function(x) summarise_enrichement(x, "dng_up", FDR = 0.05)))
M.dng.down <- do.call("rbind", lapply(enrichement_M, function(x) summarise_enrichement(x, "dng_down", FDR = 0.05)))
M.dng <- rbind(M.dng.up, M.dng.down)
```

## Females

```{r}
# Summarise BP (Biological Process) enrichment
F.bp.up <- do.call("rbind", lapply(enrichement_Fe, function(x) summarise_enrichement(x, "BP_up", FDR = 0.05)))
F.bp.down <- do.call("rbind", lapply(enrichement_Fe, function(x) summarise_enrichement(x, "BP_down", FDR = 0.05)))
F.bp <- rbind(F.bp.up, F.bp.down)

# Summarise MF (Molecular Function) enrichment
F.mf.up <- do.call("rbind", lapply(enrichement_Fe, function(x) summarise_enrichement(x, "MF_up", FDR = 0.05)))
F.mf.down <- do.call("rbind", lapply(enrichement_Fe, function(x) summarise_enrichement(x, "MF_down", FDR = 0.05)))
F.mf <- rbind(F.mf.up, F.mf.down)

# Summarise CC (Cellular Component) enrichment
F.cc.up <- do.call("rbind", lapply(enrichement_Fe, function(x) summarise_enrichement(x, "CC_up", FDR = 0.05)))
F.cc.down <- do.call("rbind", lapply(enrichement_Fe, function(x) summarise_enrichement(x, "CC_down", FDR = 0.05)))
F.cc <- rbind(F.cc.up, F.cc.down)

# Summarise KEGG pathway enrichment
F.kegg.up <- do.call("rbind", lapply(enrichement_Fe, function(x) summarise_enrichement(x, "kegg_up", FDR = 0.05)))
F.kegg.down <- do.call("rbind", lapply(enrichement_Fe, function(x) summarise_enrichement(x, "kegg_down", FDR = 0.05)))
F.kegg <- rbind(F.kegg.up, F.kegg.down)

# Summarise DO (Disease Ontology) enrichment
F.do.up <- do.call("rbind", lapply(enrichement_Fe, function(x) summarise_enrichement(x, "do_up", FDR = 0.05)))
F.do.down <- do.call("rbind", lapply(enrichement_Fe, function(x) summarise_enrichement(x, "do_down", FDR = 0.05)))
F.do <- rbind(F.do.up, F.do.down)

# Summarise DNG (Drug Ontology) enrichment
F.dng.up <- do.call("rbind", lapply(enrichement_Fe, function(x) summarise_enrichement(x, "dng_up", FDR = 0.05)))
F.dng.down <- do.call("rbind", lapply(enrichement_Fe, function(x) summarise_enrichement(x, "dng_down", FDR = 0.05)))
F.dng <- rbind(F.dng.up, F.dng.down)
```

# Summarise the results with a table

## Males

Note: There no enrichements here

```{r fig.height=10, fig.width=15}
# Biological process 
# celltypes <- do.call("c", lapply(row.names(M.bp), function(x) str_split(x, "\\.")[[1]][1]))
# M.bp$ct <- celltypes
# 
# # p1 <- ggplot(M.bp, aes(y = ct, x = Description, color = ontology)) +
# #   geom_point(size = 4) +
# #   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
# #   theme_dose(10) +
# #   xlab("") + ylab("") + 
# #   theme(panel.background = element_blank(),
# #         panel.grid.major = element_blank(), 
# #         panel.grid.minor = element_blank(),
# #         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
# #         axis.text.y = element_text(size = 16), 
# #         legend.title = element_blank(), 
# #         legend.text = element_text(size = 22), 
# #         legend.position = "top", 
# #         legend.direction="horizontal")
# 
# # MOlecular function
# celltypes <- do.call("c", lapply(row.names(M.mf), function(x) str_split(x, "\\.")[[1]][1]))
# M.mf$ct <- celltypes
# 
# p2 <- ggplot(M.mf, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") + 
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
#         axis.text.y = element_text(size = 16), 
#         legend.title = element_blank(), 
#         legend.text = element_text(size = 22), 
#         legend.position = "top", 
#         legend.direction="horizontal")
# 
# 
# # Cellular component 
# celltypes <- do.call("c", lapply(row.names(M.cc), function(x) str_split(x, "\\.")[[1]][1]))
# M.cc$ct <- celltypes
# 
# p3 <- ggplot(M.cc, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") + 
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
#         axis.text.y = element_text(size = 16), 
#         legend.title = element_blank(), 
#         legend.text = element_text(size = 22), 
#         legend.position = "top", 
#         legend.direction="horizontal")
# 
# # KEGG
# celltypes <- do.call("c", lapply(row.names(M.kegg), function(x) str_split(x, "\\.")[[1]][1]))
# M.kegg$ct <- celltypes
# 
# p4 <- ggplot(M.kegg, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") + 
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
#         axis.text.y = element_text(size = 16), 
#         legend.title = element_blank(), 
#         legend.text = element_text(size = 22), 
#         legend.position = "top", 
#         legend.direction="horizontal")
# 
# # Disease Ontology
# celltypes <- do.call("c", lapply(row.names(M.do), function(x) str_split(x, "\\.")[[1]][1]))
# M.do$ct <- celltypes
# 
# p5 <- ggplot(M.do, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") +
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1),
#         axis.text.y = element_text(size = 16),
#         legend.title = element_blank(),
#         legend.text = element_text(size = 22),
#         legend.position = "top",
#         legend.direction="horizontal")
# 
# # Disease Network
# celltypes <- do.call("c", lapply(row.names(M.dng), function(x) str_split(x, "\\.")[[1]][1]))
# M.dng$ct <- celltypes
# 
# p6 <- ggplot(M.dng, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") + 
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
#         axis.text.y = element_text(size = 16), 
#         legend.title = element_blank(), 
#         legend.text = element_text(size = 22), 
#         legend.position = "top", 
#         legend.direction="horizontal")
# 
# 
# 
# 
# print(p1)
# print(p2)
# print(p3)
# print(p4)
# print(p5)
# print(p6)
```

No Plots here, since there is no enrichment

## Female

```{r fig.height=10, fig.width=15}
# Biological process 
# celltypes <- do.call("c", lapply(row.names(F.bp), function(x) str_split(x, "\\.")[[1]][1]))
# F.bp$ct <- celltypes
# 
# p1 <- ggplot(F.bp, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") +
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1),
#         axis.text.y = element_text(size = 16),
#         legend.title = element_blank(),
#         legend.text = element_text(size = 22),
#         legend.position = "top",
#         legend.direction="horizontal")
# 
# # MOlecular function
# celltypes <- do.call("c", lapply(row.names(F.mf), function(x) str_split(x, "\\.")[[1]][1]))
# F.mf$ct <- celltypes
# 
# p2 <- ggplot(F.mf, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") + 
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
#         axis.text.y = element_text(size = 16), 
#         legend.title = element_blank(), 
#         legend.text = element_text(size = 22), 
#         legend.position = "top", 
#         legend.direction="horizontal")
# 
# 
# # Cellular component 
# celltypes <- do.call("c", lapply(row.names(F.cc), function(x) str_split(x, "\\.")[[1]][1]))
# F.cc$ct <- celltypes
# 
# p3 <- ggplot(F.cc, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") + 
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
#         axis.text.y = element_text(size = 16), 
#         legend.title = element_blank(), 
#         legend.text = element_text(size = 22), 
#         legend.position = "top", 
#         legend.direction="horizontal")
# 
# # KEGG
# celltypes <- do.call("c", lapply(row.names(F.kegg), function(x) str_split(x, "\\.")[[1]][1]))
# F.kegg$ct <- celltypes
# 
# p4 <- ggplot(F.kegg, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") + 
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1), 
#         axis.text.y = element_text(size = 16), 
#         legend.title = element_blank(), 
#         legend.text = element_text(size = 22), 
#         legend.position = "top", 
#         legend.direction="horizontal")
# 
# # Disease Ontology
# celltypes <- do.call("c", lapply(row.names(F.do), function(x) str_split(x, "\\.")[[1]][1]))
# F.do$ct <- celltypes
# # 
# p5 <- ggplot(F.do, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") +
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1),
#         axis.text.y = element_text(size = 16),
#         legend.title = element_blank(),
#         legend.text = element_text(size = 22),
#         legend.position = "top",
#         legend.direction="horizontal")
# 
# # Disease Network
# celltypes <- do.call("c", lapply(row.names(F.dng), function(x) str_split(x, "\\.")[[1]][1]))
# F.dng$ct <- celltypes
# 
# p6 <- ggplot(F.dng, aes(y = ct, x = Description, color = ontology)) +
#   geom_point(size = 4) +
#   scale_color_manual(name = "Gene Set", values = safe_colorblind_palette[c(2,1)]) +
#   theme_dose(10) +
#   xlab("") + ylab("") +
#   theme(panel.background = element_blank(),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.text.x = element_text(size = 14, angle = 80, hjust = 1, vjust = 1),
#         axis.text.y = element_text(size = 16),
#         legend.title = element_blank(),
#         legend.text = element_text(size = 22),
#         legend.position = "top",
#         legend.direction="horizontal")
# 
# 
# 
# 
# print(p1)
# print(p2)
# print(p3)
# print(p4)
# print(p5)
# print(p6)
```

Same for females